Follow-up Samples

### 2016-12-09 submission/manuscript_outline_20161024.tex
update abstract and Importance blurb

Colorectal cancer (CRC) continues to be a worldwide health problem with
early detection being used as a key component in mitigating deaths due
to the disease. Previous research suggests that the bacterial microbiome
can be used as a biomarker for CRC. In this study, we used a
microbiome-based model that classified individuals as having an adenoma
or carcinoma to assess whether the probability of having one of these
lesions changed after surgical treatment. We also characterized the
change in the gut microbiota before and after the surgical treatment.
This model was tested on a 66 person group that included samples before
and after treatment to allow for the assessment of how the model adjusts
risk after treatment. The model used for prediction had an AUC of 0.813.
For the follow up samples our Random Forest model significantly
decreased the positive probability of lesion compared to the initial
samples for both adenoma (P-value = 7.95e-08) and carcinoma (P-value =
7.95e-08). Our model predicted that 36.4\% of the 67-person cohort had
normal colons and 63.6\% had a lesion. Some OTUs that changed the most
before and after treatment included OTUs that were affiliated with
members of Blautia, Clostridium\_XlVa and Escherichia/Shigella. Our
model suggests that treatment does significantly reduce the probability
of having a colonic lesion. Further surveillance of these individuals
will enable us to determine whether models such as the one we present
here can also be used to predict recurrence of colorectal cancer.

@@

This is one of the first studies to investigate within humans what
happens to the bacterial microbiome before and after colorectal cancer
treatment. Specifically, it aims to assess how the random forest machine
learning algorithm built models respond to treatment and adjust their
positive probability calls of whether the individual has an adenoma or
carcinoma.

>>>>

Colorectal cancer (CRC) continues to be a worldwide health problem with
early detection being used as a key component in mitigating deaths due
to the disease. Previous research suggests a link between stool
bacterial microbiome and CRC. In this study, we used a model based on
the microbiome, demographics, and prior medical history to classify
individuals as having a lesion (i.e.~adenoma or carcinoma). We then used
this model to characterize the change in the gut microbiota before and
after surgery. The overall objective was to investigate the changes in
the microbiome after surgery in patients with adenomas or carcinomas.
This model was tested on a 66 person group that included samples before
and after treatment to allow for the assessment of how the model adjusts
risk after treatment. The model used for prediction had an AUC of 0.813.
For the follow up samples our Random Forest model significantly
decreased the positive probability of lesion compared to the initial
samples for both adenoma (P-value = 7.95e-08) and carcinoma (P-value =
7.95e-08). Our model predicted that 36.4\% of the 67-person cohort had
normal colons and 63.6\% had a lesion. Some OTUs that changed the most
before and after treatment included OTUs that were affiliated with
members of Blautia, Clostridium\_XlVa and Escherichia/Shigella. Our
model suggests that treatment does significantly reduce the probability
of having a colonic lesion. Further surveillance of these individuals
will enable us to determine whether models such as the one we present
here can also be used to predict recurrence of colorectal cancer.

@@

This is one of the first studies to investigate within humans what
happens to the bacterial microbiome before and after adenoma and
carcinoma treatment. Specifically, it aims to assess how a random forest
machine learning algorithm built model respond to treatment and adjust
it's positive probability calls of whether the individual has an adenoma
or carcinoma due to surgical removal of the lesion.
